### 2019-02-01 Low_latency_control/images/multiplex.tex

\begin{tikzpicture}
\draw[step=.5cm,gray,thick] (-2,-2) grid (3,2);
\node (u1) at (-2,-2) {};
\draw [ultra thick, draw=black, fill=black!20!white] (-2,-2) rectangle  (0,0);
\node (u2) at (-2,0) {};
\draw [ultra thick, draw=black, fill=red!20!white] (-2,0)  rectangle  (0,1);
\draw [
    thick,
    decoration={
        brace,
        raise=0.1cm
    },
    decorate
] (u1.west) -- (u2.west) 
node [pos=0.5,anchor=north,rotate=90,yshift=.7cm] {RU};
\draw [ultra thick, draw=black, fill=blue!20!white] (-2,1)  rectangle  (0,2);
\draw [ultra thick, draw=black, fill=green!20!white] (0,-2) rectangle (1,2);
\draw [ultra thick, draw=black, fill=cyan!20!white] (1,-2)  rectangle  (3,0);
\draw [ultra thick, draw=black, fill=magenta!20!white] (1,0)  rectangle  (3,2);
\node[align=center, scale =0.9] at (0,-2.5) {Transmission time};
%\node[rotate=90] at (-2.5,0) {Resource units (RU)};
\end{tikzpicture}