### 2019-02-01 Low_latency_control/images/tikz_random_access.tex

%!TEX root = root.tex
\documentclass[]{standalone}

\usepackage{amsmath} % nice math symbols

\usepackage{color} % change text color        
\usepackage{tikz}
\usetikzlibrary{matrix} % for block alignment
\usetikzlibrary{arrows, decorations.markings} % for arrow heads
\usetikzlibrary{calc} % for manimulation of coordinates
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{shapes}

\usetikzlibrary{fit}

%\usetikzlibrary{backgrounds}

\begin{document}
% TikZ styles for drawing

\pgfdeclarelayer{bg0}    % declare background layer
\pgfdeclarelayer{bg1}    % declare background layer
\pgfsetlayers{bg0,bg1,main}  % set the order of the layers (main is the standard layer)

\tikzstyle{block} = [draw,rectangle,thick,
%minimum height=0.7cm, minimum width=0.3cm, 
text height=0.2cm, text width=0.7cm, 
fill=blue!30, outer sep=0pt, inner sep=0pt]
\tikzstyle{dots} = [font = \large, minimum width=2pt]
\tikzstyle{dash_block} = [draw,rectangle,dashed,minimum height=1cm,minimum width=1cm]
\tikzstyle{smallblock} = [draw,rectangle,minimum height=0.5cm,minimum width=0.5cm,fill= blue!30, font =  \scriptsize]
\tikzstyle{smallcircle} = [draw,ellipse,minimum height=0.1cm,minimum width=0.3cm,fill= yellow!40, font =  \scriptsize ]
\tikzstyle{connector} = [->]
\tikzstyle{dash_connector} = [->,thick,decorate,decoration={snake, amplitude =1pt, segment length=8pt}, magenta]
\tikzstyle{branch} = [circle,inner sep=0pt,minimum size=1mm,fill=black,draw=black]

\tikzstyle{vecArrow} = [thick, decoration={markings,mark=at position
   1 with {\arrow[semithick]{open triangle 60}}},
   double distance=1.4pt, shorten >= 5.5pt,
   preaction = {decorate},
   postaction = {draw,line width=1.4pt, white,shorten >= 4.5pt}]

\tikzset{>=latex}

\begin{tikzpicture}[scale=1, blocka/.style ={rectangle,text width=0.6cm,text height=0.5cm, outer sep=0pt}]
 \small

    % node placement with matrix library: 5x4 array
    \matrix(M)[ampersand replacement=\&, row sep=0.25cm, column sep=0.25cm] {

    \node[smallblock, align=center, fill= green!30] (P1) {Plant {1}};\&

		\node[smallblock, align=center] (S1) {Sensor {1}};\&
		\&
		\&
		\node[smallblock, align=center, fill= green!30] (P2) {Plant {2}};\&

		\node[smallblock, align=center] (S2) {Sensor {2}};\&

%    \&
    %\node(d0) {$\cdots$};\&
%    \&

    \\

    %\node(d1) {$\cdots$};\&
%    \&

    \\
		\\
		\\
		\\
		\\
		\\
		\\

		\&
		\node[smallblock, align=center] (C1) {Controller {1}};\&
		\&
		\&
		\&
		\node[smallblock, align=center] (C2) {Controller {2}};\&

		\\
    %%
    %\node[blocka] (R1) {};\&
%%    \node[smallcircle] (R2) {R2};\&
    %%\node[blocka] (d3) {};\&
    %\node[blocka] (R2) {};\&
    %\\
    };

    %\node[block] (outer) [fit=(R1.north west)  (R2.south east)] {};
    %
    %\node[align=center, font =  \scriptsize] at (outer.center) {Access Point/ \\Controller};

    \draw [->,  red] (S1.south) -- node[left, pos = 0.1]{ $\alpha_1$ } (C1.north);
%    \draw [->, thick, magenta] (T2) -- node[left]{ \scriptsize $h_2$} (R2);
    \draw [->, red] (S2.south) -- node[right, pos = 0.1]{ $\alpha_2$ } (C2.north);
%

		\draw [->, dashed,  red] (S1.south) -- node[right, pos=0.2]{ } (C2.north);

		\draw [->, dashed, red] (S2.south) -- node[left, pos=0.2]{ } (C1.north);

		%\begin{pgfonlayer}{bg0}    % select the background layer
		%\draw [->, black] (C1.west) |- ($(C1) + (-25pt, 0pt)$) node(down_right){} 
		%-- ($(P1) + (-25pt,+0pt)$) node(up_right){} -- (P1);
		%\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}    % select the background layer
		\draw [->, black] (C1.west) -|   (P1.south);
		\end{pgfonlayer}

		%\begin{pgfonlayer}{bg0}
		%\draw [->,  black] (R2) |- ($(R2) + (-25pt,-15pt)$) node(down_right){} 
		%--($(P2) + (-25pt,+0pt)$) node(up_right){} -- (P2);
		%\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}    % select the background layer
		\draw [->, black] (C2.west) -|   (P2.south);
		\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}
		\draw [->, black] (P1) -- (S1);
		\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}
		\draw [->,  black] (P2) -- (S2);
		\end{pgfonlayer}

		%
		\begin{pgfonlayer}{bg1}
		%\begin{scope}[on background layer]
		\node(shared) [fill=red!10, fit={
		($(S1.south) + (-32pt, -12pt)$) 
		($(S2.south) + (+8pt, -12pt)$)
		($(C1.north) + (-32pt, +10pt)$)
		($(C2.north) + (+8pt, +10pt)$)
		}] {};
		%\end{scope}
		\end{pgfonlayer}

		\node[align=center, red!40, font =  \footnotesize](shared_medium) at ($(shared.center)-(50pt, 0)$) {Shared \\ Wireless \\ Medium};

\coordinate (FIRST NE) at (current bounding box.north east);
   \coordinate (FIRST SW) at (current bounding box.south west);

	\pgfresetboundingbox
   \useasboundingbox ($(FIRST SW) + (+52pt,+3pt)$) rectangle ($(FIRST NE) + (-10pt,-2pt)$);

\end{tikzpicture}
\end{document}